### 2017-12-08 1/exams/f-practice.tex
Practice final.

\begin{center}
  {\Large {\bf CMSC 131A, Midterm 2}\\
    \censor{{\bf SOLUTION}} \ \\
    \ \\
    Fall 2017 \ \\
  }
\end{center}

@@

\begin{center}
\begin{tabular}{| c | c |}
  \hline
  Question & Points \\ \hline \hline
  1 & 15 \\ \hline
  2 & 12 \\ \hline
  3 & 20 \\ \hline
  4 & 15 \\ \hline
  5 & 15 \\ \hline
  6 & 20 \\ \hline\hline
  Total: & 97\\
  \hline
\end{tabular}
\end{center}

@@

\noindent
{\bf Problem 1 (15 points).}
%
Here are some data definition relevant to representing a dictionary,
which associates words with the their definitions:

\begin{verbatim}
;; A Dict is one of:
;; - '()
;; - (cons (list String String) Dict)
;; Interp: a collection of definitions where each element is a 
;; two-element list of a word (first) and its meaning (second).

;; A Result is one of:
;; - #false
;; - String
;; Interp: represents the result of looking up a word in the
;; dictionary; #false means not found.
\end{verbatim}

@@

\begin{verbatim}
;; lookup : String Dict -> Result
;; Lookup the (first) definition of a word in the dictionary,
;; produces #false if not found
\end{verbatim}

@@

(check-expect (lookup "a" '()) #false)
(check-expect (lookup "a" (list (list "a" "b"))) "b")
(check-expect (lookup "a" (list (list "c" "b") (list "a" "d"))) "d")
(define (lookup w d)
  (cond [(empty? d) #false]
        [(cons? d)
         (if (string=? (first (first d)) w)
             (second (first d))
             (lookup w (rest d)))]))

@@

\noindent
{\bf Problem 2 (12 points).}
%
For each of the following functions, provide the most general signature
that correctly describes the function:

\begin{alltt}
;; method : \censor{Number -> Number}

@@

;; raekwon : \censor{[Listof Number] -> Boolean}

@@

;; rza : \censor{[Listof String] [Listof Number] -> [Listof Number]}

@@

;; gza : \censor{(String Number -> Number) -> Number}

@@

;; odb : \censor{(Number -> Number) -> (Number -> Number)}

@@

;; ghostface : \censor{[X Y] [Listof X] [X -> Y] -> [Listof Y]}

@@

\noindent
{\bf Problem 3 (20 points).}
%
Design a program called {\tt w-avg} that computes the weighted average
of a list of numbers and a list of weights.

For example, let's say a class grade is based on two midterms and a
project where the project is worth twice as much as the midterms.  A
student who gets an 80 and a 70 on the midterms and a 90 on the
project would have a weighted average of 82.5: $((1 \times 80) + (1
\times 70) + (2 \times 90)) / 4$, which can be computed with {\tt
  (w-avg (list 80 70 90) (list 1 1 2))}.  You should assume the two
lists have the same length.

@@

;; w-avg : [Listof Number] [Listof Number] -> Number
;; Compute the weighted average of a list of numbers and weights
;; Assume: lists have the same length (and non-empty)
(check-expect (w-avg (list 70 80 90) (list 1 1 2)) 82.5)
(define (w-avg lon ws)
  (/ (w-sum lon ws) 
     (foldr + 0 ws)))

;; w-sum : [Listof Number] [Listof Number] -> Number
;; Sum the list of numbers according to given weights
;; Assume: lists have same length
(check-expect (w-sum (list 70 80 90) (list 1 1 2)) (+ 70 80 180))
(define (w-sum lon ws)
  (cond [(empty? lon) 0]
        [(cons? lon)
         (+ (* (first lon) (first ws))
            (w-sum (rest lon) (rest ws)))]))

@@

\newpage 
\noindent
{\bf Problem 4 (15 points).}
%
Here is a parametric data definition for a tree of elements:
\begin{verbatim}
;; A [Tree X] is one of:
;; - (make-leaf)
;; - (make-node X [Tree X] [Tree X])
;; Interp: a binary tree that is either empty (a leaf), or non-empty (a node)
;; with an element and two sub-trees.
(define-struct leaf ())
(define-struct node (elem left right))
\end{verbatim}
%% {\bf Problem 4(a) (10 points).} Design a function, {\tt tree-sum},
%% that operates on a binary tree of numbers, {\tt [Tree Number]}, and
%% produces the sum of all the numbers in the tree.

%% \begin{SaveVerbatim}{VerbEnv}

%% SOLUTION:

%% ;; tree-sum : [Tree Number] -> Number
%% ;; Sum all the numbers in the tree
%% (check-expect (tree-sum (make-leaf)) 0)
%% (check-expect (tree-sum (make-node 7 (make-node 1 (make-leaf) (make-leaf)) 
%%                                      (make-leaf)))
%%               8)
%% (define (tree-sum bt)
%%   (cond [(leaf? bt) 0]
%%         [(node? bt)
%%          (+ (node-elem bt)
%%             (tree-sum (node-left bt))
%%             (tree-sum (node-right bt)))]))

%% \end{SaveVerbatim}

%% \censor{%
%% \noindent
%% \BUseVerbatim{VerbEnv}}

%% \newpage

\noindent
%{\bf Problem 4(b) (10 points).}
Here is a function that produces a list of all the elements in a tree:
\begin{verbatim}
;; tree-elems : [X] . [Tree X] -> [Listof X]
;; Produce a list of all elements in tree going top-down, left-to-right
(check-expect (tree-elems (make-leaf)) '())
(check-expect (tree-elems (make-node 7 
                                     (make-node 9 (make-leaf) (make-leaf))
                                     (make-node 2 (make-leaf) (make-leaf))))
              (list 7 9 2))
(define (tree-elems bt)
  (cond [(leaf? bt) '()]
        [(node? bt)
         (cons (node-elem bt)
               (append (tree-elems (node-left bt))
                       (tree-elems (node-right bt))))]))
\end{verbatim}
Here is an abstraction function for trees that is similar to
{\tt foldr} for lists, but works on trees:
\begin{verbatim}
;; tree-fold : [X Y] . [X Y Y -> Y] Y [Tree X] -> Y
;; The fundamental abstraction function for trees
(define (tree-fold f b bt)
  (cond [(leaf? bt) b]
        [(node? bt)
         (f (node-elem bt)
            (tree-fold f b (node-left bt))
            (tree-fold f b (node-right bt)))]))
\end{verbatim}
Give an equivalent definition of {\tt tree-elems} in terms of {\tt
  tree-fold}.  (Just provide the code.)

@@

\noindent
[Space for problem 4.]

\newpage
\noindent
{\bf Problem 5 (15 points).}  
%
Design a program that takes a list of strings and produces the count
of strings with length longer than 3 in the list.  For example, if the
list contains {\tt "a"}, {\tt "dave"}, {\tt "abc"}, and {\tt "fred"},
the count is 2.  You may assume the {\tt{[Listof String]}} data
definition is defined.
%
For full credit, use list abstraction functions.  For partial credit,
follow the template for {\tt{[Listof String]}}.

;; count-3+ : [Listof String] -> Number
;; Count the number of strings with length more than 3 in given list
(check-expect (count-3+ '()) 0)
(check-expect (count-3+ (list "a" "abcd" "fred" "c")) 2)
(define (count-3+ los)
  (foldr (lambda (s c) (if (> (string-length s) 3) (add1 c) c)) 0 los))

;; Alts:
(define (count-3+ los)
  (local [;; count+ : String Number -> Number
          ;; Bump count if string has length > 3
          (define (count+ s c) 
            (if (> (string-length s) 3) (add1 c) c))]
    (foldr count+ 0 los))

(define (count-3+ los)
  (local [;; 3+? : String -> Boolean
          ;; Does the string have length > 3?
          (define (3+? s) 
            (> (string-length s) 3))]
    (length (filter 3+? los))))

(define (count-3+ los)
  (length (filter (lambda (s) (> (string-length s) 3)) los)))
\end{SaveVerbatim}

@@

\noindent
{\bf Problem 6 (20 points).}
%
Here's a data definition for choose-your-own-adventure books:
\begin{verbatim}
;; A CYOA is one:
;; - String
;; - (make-choice String CYOA CYOA)
(define-struct choice (q yes no))
;; Interp: a choose-your-own-adventure book where a string
;; is a conclusion and a choice is a yes/no question and
;; the two adventures that follow as a consequnce of the answer.
\end{verbatim}

\noindent
Design a function called {\tt ans} that consumes a {\tt CYOA} and a
{\tt [Listof Boolean]} that represents answers to the yes/no
questions: {\tt \#true} means yes, {\tt \#false} means no.  The {\tt
  ans} function produces the {\tt CYOA} adventure that results after
giving all of the answers in the list.  If the list contains too many
answers, {\tt ans} just produces the conclusion of the adventure.

(define a0 "A0")
(define a1 (make-choice "Q1" "A1" "A2"))
(define a2 (make-choice "Q2" a0 a1))

;; ans : CYOA [Listof Boolean] -> CYOA
;; Compute the remaining adventure after using answers from list
(check-expect (adv a0 (list #true #false)) "A0")
(check-expect (adv a1 (list #true #false)) "A1")
(check-expect (adv a2 (list #false #true)) "A1")
(check-expect (adv a2 (list #false)) a1)
(define (ans adv loa)
  (cond [(string? adv) adv]
        [(empty? loa) adv]
        [(cons? loa)
         (if (first loa)
             (ans (choice-yes adv) (rest adv))
             (ans (choice-yes adv) (rest adv)))]))
\end{SaveVerbatim}

\end{document}

>>>>

\begin{center}
  {\Large {\bf CMSC 131A, Final Exam (Practice)}\\
    \censor{{\bf SOLUTION}} \ \\
    \ \\
    Fall 2017 \ \\
  }
\end{center}

@@

\begin{center}
\begin{tabular}{| c | c |}
  \hline
  Question & Points \\ \hline \hline
  1 & 10 \\ \hline
  2 & 10 \\ \hline
  3 & 15 \\ \hline
  4 & 12 \\ \hline
  5 & 15 \\ \hline
  6 & 15 \\ \hline
  7 & 15 \\ \hline
  8 & 20 \\ \hline\hline
  Total: & 112\\
  \hline
\end{tabular}
\end{center}

@@

\noindent
{\bf Problem 1 (10 points).}
%
For the following program, write out each step of computation. At each
step, underline the expression being simplified. Label each step as
being ``arithmetic'' (meaning any built-in operation),
``conditional'', ``plug'' (for plugging in an argument for a function
parameter), or ``constant'' for replacing a constant with its value.

\begin{verbatim}
(define T 7)
(define (q z) (sqr z))
(cond [(> T 3) (q 4)]
      [else 9])
\end{verbatim}

\noindent
{\bf Problem 2 (10 points).}
%
For the following structure definition, list the names of every
function it creates.  For each function, classify it as being either a
constructor, accessor, or predicate.

\begin{verbatim}
(define-struct zip (pad msg))
\end{verbatim}

\begin{SaveVerbatim}{VerbEnv}  
SOLUTION:

- make-zip  : Constructor
- zip-pad  : Accessor
- zip-msg   : Accessor
- zip?      : Predicate
\end{SaveVerbatim}

\noindent
{\bf Problem 3 (15 points).}
%
Here are some data definition relevant to representing a dictionary,
which associates words with the their definitions:

\begin{verbatim}
;; A Dict is one of:
;; - '()
;; - (cons (list String String) Dict)
;; Interp: a collection of definitions where each element is a 
;; two-element list of a word (first) and its meaning (second).
\end{verbatim}

@@

\begin{verbatim}
;; update : String String Dict -> Dict
;; Update entry for given word and meaning, if it exists
;; Add the entry if it does not
\end{verbatim}

@@

(check-expect (update "a" "b" '()) (list (list "a" "b")))
(check-expect (update "a" "c" (list (list "a" "b")))
              (list (list "a" "c")))
(check-expect (update "a" "e" (list (list "c" "b") (list "a" "d"))) 
              (list (list "c" "b") (list "a" "e")))
(define (update w m d)
  (cond [(empty? d) (list (list w m))]
        [(cons? d)
         (if (string=? (first (first d)) w)
             (cons (list w m) d)
             (cons (first d) (lookup w (rest d))))]))

@@

\noindent
{\bf Problem 4 (12 points).}
%
For each of the following functions, provide the most general {\bf type} signature
in Typed Racket that correctly describes the function:

\begin{alltt}
(: method : \censor{Number -> Number                                    })

@@

(: raekwon : \censor{[Listof Number] -> Boolean                         })

@@

(: rza : \censor{[Listof String] [Listof Number] -> [Listof Number]     })

@@

(: gza : \censor{(String Number -> Number) -> Number                    })

@@

(: odb : \censor{(Number -> Number) -> (Number -> Number)               })

@@

(: ghostface : \censor{[X Y] [Listof X] [X -> Y] -> [Listof Y]          })

@@

\noindent
{\bf Problem 5 (15 points).}
%
Here is the design of a program {\tt w-avg} that computes the weighted
average of a list of numbers and a list of weights.

Design an accumulator variant of {\tt w-avg} that maintains two
accumulators: the weighted sum of scores and the sum of weights.
Redefine {\tt w-avg} in terms of your accumulator design.

For example, let's say a class grade is based on two midterms and a
project where the project is worth twice as much as the midterms.  A
student who gets an 80 and a 70 on the midterms and a 90 on the
project would have a weighted average of 82.5: $((1 \times 80) + (1
\times 70) + (2 \times 90)) / 4$, which can be computed with {\tt
  (w-avg (list 80 70 90) (list 1 1 2))}.

In the accumulator version, after all elements of the list have been
seen, the first accumulator would be $330 = 80+70+2\times 90$) and
the second accumulator would be $4 =1+1+2$.

\begin{verbatim}
;; w-avg : [Listof Number] [Listof Number] -> Number
;; Compute the weighted average of a list of numbers and weights
;; Assume: lists have the same length (and non-empty)
(check-expect (w-avg (list 70 80 90) (list 1 1 2)) 82.5)
(define (w-avg lon ws)
  (/ (w-sum lon ws) 
     (foldr + 0 ws)))

;; w-sum : [Listof Number] [Listof Number] -> Number
;; Sum the list of numbers according to given weights
;; Assume: lists have same length
(check-expect (w-sum (list 70 80 90) (list 1 1 2)) (+ 70 80 180))
(define (w-sum lon ws)
  (cond [(empty? lon) 0]
        [(cons? lon)
         (+ (* (first lon) (first ws))
            (w-sum (rest lon) (rest ws)))]))
\end{verbatim}

\noindent
Your solution should not need a helper function like {\tt w-sum}.

@@

@@

[Space for problem 5.]

\newpage 
\noindent
{\bf Problem 6 (15 points).}
%
Here is a parametric data definition for a tree of elements:
\begin{verbatim}
;; A [Tree X] is one of:
;; - (make-leaf)
;; - (make-node X [Tree X] [Tree X])
;; Interp: a binary tree that is either empty (a leaf), or non-empty (a node)
;; with an element and two sub-trees.
(define-struct leaf ())
(define-struct node (elem left right))
\end{verbatim}

\noindent
Here is a function that counts the elements in a tree:
\begin{verbatim}
;; tree-count : [X] . [Tree X] -> Natural
;; Count all the elements in a tree
(check-expect (tree-count (make-leaf)) 0)
(check-expect (tree-count (make-node 7 
                                     (make-node 9 (make-leaf) (make-leaf))
                                     (make-node 2 (make-leaf) (make-leaf))))
              3)
(define (tree-count bt)
  (cond [(leaf? bt) 0]
        [(node? bt)
         (add1 (+ (tree-count (node-left bt))
                  (tree-count (node-right bt))))]))
\end{verbatim}
Here is an abstraction function for trees that is similar to
{\tt foldr} for lists, but works on trees:
\begin{verbatim}
;; tree-fold : [X Y] . [X Y Y -> Y] Y [Tree X] -> Y
;; The fundamental abstraction function for trees
(define (tree-fold f b bt)
  (cond [(leaf? bt) b]
        [(node? bt)
         (f (node-elem bt)
            (tree-fold f b (node-left bt))
            (tree-fold f b (node-right bt)))]))
\end{verbatim}
Give an equivalent definition of {\tt tree-count} in terms of {\tt
  tree-fold}.  (Just provide the code.)

@@

\noindent
[Space for problem 6.]

\newpage
\noindent
{\bf Problem 7 (15 points).}  
%
During your internship at the University of Maryland's Center for
Bioinformatics and Computational Biology (CBCB), you're asked to
design a program for transcribing DNA strands into RNA strands.
%
A DNA strand is an arbitrarily long sequence of bases, which are
either: C, G, A, or T; an RNA strand is an arbitrarily long sequences
of either: C, G, A, or U.  The \emph{transcription} of DNA to RNA
replaces each occurrence of T with U.

Design a program that, given a DNA strand, computes its transcription
to RNA.

@@

\noindent
{\bf Problem 8 (20 points).} Here is the data definition for graphs we studied in class:
\begin{verbatim}
;; A Graph is [Listof (cons String [Listof String])]
;; Interp: a graph is a list of nodes and neighbors reachable from the node

;; A Path is a [Listof String]
;; Interp: series of nodes neighboring each other that connect a source
;; to destination in some graph.
\end{verbatim}

\begin{verbatim}
;; all-paths : Graph String String -> [Listof Path]
;; Produce all non-cyclic paths from src to dst in g.
;; Assume: both nodes exist in graph.
(define (all-paths g src dst) '())
\end{verbatim}

\noindent
You may assume the following function is already defined and works correctly:

\begin{verbatim}
;; neighbors : Graph String -> [Listof String]
;; Produce the list of immediate neighbors of given node name.
;; Assume: node exists in graph.
\end{verbatim}

\noindent
Note: this function is tricky to test since the tests shouldn't
specify the order of the paths in the list.  You may ignore this
subtly and just write tests as though a {\tt check-expect} will pass
if the checked and expected expressions produce lists with the same
elements in any order, e.g. {\tt (check-expect (list 1 2) (list 2 1))}
would pass.

\newpage
[Space for problem 8.]
\end{document}
